\chapter{Model Systems with Stochasticity}
So far: 
$$\dot{\myM{x}}=f(\myM{x},\myM{u}).$$
Now: 
$$\dot{\myM{x}}=f(\myM{x},\myM{u},\myM{w}),$$
where $\myM{w}$ is the output of some random process. Here we only cover \textbf{process noise}, caused by:
\begin{itemize}
\item Disturbances
\item Model Uncertainty
\end{itemize}
Measurement noise will not be considered here.

Every time that we've analyzed a system to date, we've asked questions like "given x[0], what is the long-term behavior of the system? But now $x[n]$ is a random variable. The trajectories of this system do not converge, and the system does not exhibit any form of stability that we've introduced so far. 


All is not lost. The \textbf{distribution} of this random variable is actually very well behaved. This is the key idea for this chapter. Let us use $p_n(x)$ to denote the probability density function over the random variable $\myM{x}$ at time $n$.

It is actually possible to write the dynamics of the probability density with the simple relation 
$$p_{n+1}(x) = \int_{-\infty}^\infty p(x|x') p_n(x') dx',$$
where $p(x|x')$ encodes the stochastic dynamics as a conditional distribution of the next state (here x) as a function of the current state (here $x'$). Dynamical systems that can be encoded in this way are known as continuous-state Markov Processes, and the governing equation above is often referred to as the "master equation" for the stochastic process.

\section{The Master Equation}
\section{Stationary Distributions}
\section{Extended Example: The Rimless Wheel on Rough Terrain}
\section{Noise Models for Real Robots}
\section{Analysis: A Preview}
\section{Control Design: A Preview}